\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-06-25 paper/main_v1.tex
diff, original manuscript stored in main_v1

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function. CSF
provides mechanical support for the brain and helps
distribute nutrients, neurotransmitters and metabolites throughout the
central nervous system. CSF flow is driven by several processes,
including the beating of motile cilia located on the
walls of the brain ventricles. Despite the physiological importance of
CSF, the underlying mechanisms of CSF flow and solute transport
in the brain ventricles remain to be comprehensively
resolved. This study analyzes and evaluates specifically
the role of motile cilia in CSF flow and transport. 
We developed finite element methods for modeling flow
and transport using the geometry of the zebrafish larval
brain ventricles, for which we have detailed knowledge of
cilia properties and CSF motion.
The computational model is validated by in vivo experiments
that monitor transport of a photoconvertible protein secreted in the
brain ventricles. Our results show that while cilia contribute to
advection of large particles, diffusion plays a significant role in
the transport of small solutes. We also demonstrate how cilia location
and the geometry of the ventricular system impact solute
distribution. Altogether, this work presents a computational framework
that can be applied to other ventricular systems, together with new
concepts of how molecules are transported within the brain and its
ventricles.

@@

% I4: How do we address the problem and what do we find out?
In this study, we combined computational and experimental approaches to investigate
the role of motile cilia in brain ventricular CSF flow and transport.
We model the flow and transport by partial differential equations and solve
them numerically with finite element methods to compute flow patterns and
solute concentration dynamics. A geometry representing the zebrafish larval brain ventricles
was used, for which we have detailed knowledge of cilia properties and CSF motion velocities
and frequencies~\cite{Olstad2019CiliaryDevelopment}. 
Notably, this early developmental stage allows us to simplify our model
by omitting CSF secretion, which starts at later stages~\cite{Jeong2024TheZebrafish},
while still being well-conserved with mammals~\cite{DGama2025MotileBrain,
Olstad2019CiliaryDevelopment, Ringers2020Role, DGama2021Diversity, jurisch2020radial}. 
The computational model is validated against experimental data by monitoring in vivo
the dynamics of a photoconvertible fluorescent protein Dendra2
expressed within the brain ventricles.
Our results show that diffusion plays a major role in the transport of small solutes,
whereas cilia are paramount for the movement of larger particles.
We also found that the location of cilia, as well as changes to the ventricular geometry,
largely impact solute distribution.
In summary, our work presents a computational
framework that can be applied to other
ventricular systems in the future, together with new concepts
of how cilia distribute molecules and
particles within the brain ventricles.

@@

In vivo particle tracking reveals characteristic CSF flow patterns in
zebrafish larval brain ventricles
(\Crefpanels{fig:fig1}{a--c})~\cite{Olstad2019CiliaryDevelopment}.
We attempt to accurately represent these flow patterns with
computational fluid dynamics driven by cardiac pulsations and motile
cilia lining the surface of the brain ventricles.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein (\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video S2}).
To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{
    \textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$ 
    in each region of interest (ROI)
    observed during experiments monitoring
    the movement of Dendra2 proteins photoconverted by laser targeting ROI 1. 
    Bold lines are mean values of the cohort ($n=8$) and shaded regions cover
    one standard deviation.
    \textbf{b.} Mean concentration $\cbar(t)$ in each ROI in a
    simulation of Dendra2 transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F(t)$ reaches a threshold value of 0.25 in
    one representative zebrafish larva.
    \textbf{d.} Time to $\cbar(t)$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    The posterior-most regions of the geometries are excluded,
    because $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions. Scale bars 50 \textmu m.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.80, 0.76, 0.86, 0.60 and 0.38

@@

We previously characterized three distinct ciliated cell lineages in
the zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment,
  DGama2025MotileBrain}.  To identify the effects of these lineages on
CSF flow and transport, we investigate region-specific cilia paralysis
with the computational model.  Three cilia-covered regions were
paralyzed separately: the anterior ventricle, and the dorsal and
ventral parts of the middle ventricle (\Crefpanels{fig:fig5}{a}).  To
simulate regional cilia paralysis, we remove the tangential traction exerted by
the cilia in the respective region.

@@

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow
patterns, and we previously identified large
interindividual variation in ventricular size and
CSF velocities between larva at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
To assess how these
variations influence solute distribution, we consider four alternative
ventricular geometries: (i) constriction of the anterior-middle
interventricular duct (66 \% reduction in cross-sectional area);
(ii) constriction of the middle-posterior duct (33 \% reduction);
(iii) middle ventricle shrunk in the lateral direction (43 \% reduction);
and (iv) all three combined (\Crefpanels{fig:fig6}{a, b}).
In these four modified geometries, the total volume is reduced by
3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively, compared to the original
geometry. As a result of the geometry modifications,
the forces applied with the 
cilia tangential traction differ across the scenarios
(\Crefpanels{fig:fig6}{c}).

@@

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within brain ventricles.  We used the larval
zebrafish brain ventricles since their geometry and flow mechanisms are
well described, thereby allowing for validation of the simulation
results against experimental measurements. Our computational framework
can be adapted to other ventricular systems, once their flow parameters
are adequately resolved. Here, we assumed that
intraventricular CSF flow was governed by the Stokes equations. By
imposing tangential traction and normal pressure acting on the CSF, we
were able to reproduce flow features observed experimentally. In
particular, vortex-structured flow mediated by the cilia and pulsatile
flow mediated by the cardiac cycle emerged as key flow characteristics.

@@

Removing motile cilia delayed transport within
the middle ventricle and towards the anterior ventricle,
while the transport to the posterior ventricle was unaffected.
Photoconversion experiments with cilia mutant
zebrafish larva (\emph{smh}) aligned with these observations.
These simulations and experiments further indicate that
transport from the middle to the posterior ventricle is
purely diffusive at this developmental stage,
while transport within the middle ventricle
depends on cilia motility.
Furthermore, our experimental results provide quantification
of the impact of cilia motility
on brain ventricular transport.
Interestingly, Yoshida~\emph{et al.} observed a drastic reduction
in CSF mixing in specific regions of the human brain
ventricles when simulating reduced cilia motility~\cite{Yoshida2022EffectVentricles},
an observation that compares well with our findings.

@@

We modeled motile cilia as a tangential
traction at the ventricular walls
while allowing the CSF to slip,
motivated by experimental observations of
flow fields in zebrafish larval brain ventricles in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The ciliated boundaries in our model are regionally homogeneous,
whereas in reality cilia populate the ventricle walls
heterogeneously~\cite{Olstad2019CiliaryDevelopment}. Heterogeneity
in cilia location, beating direction and beating frequency
enhances particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Beating directionality can affect transport,
and specific coordination of the ciliary beating is important for
flow and signaling processes in the brain ventricles
\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
for physiological mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet},
and in the nodal flow that establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry, Sawamoto2006NewBrain}.
Incorporating spatial heterogeneity in our model would
therefore presumably affect results, and could be
interesting to pursue in further work.

@@

\subsection{Outlook}
Altogether, we have shown that the flow
generated by ciliary motion in the zebrafish larval brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In contrast, one can expect that in large brains such as the
human brain, ciliary beating contributes only to flow proximal
to the ventricular walls\cite{Siyahhan2014FlowVentricles}.
Nevertheless, the cilia may still play an important role in regulating the
amount of substances, especially extracellular vesicles,
close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Notably, the discovery of a cilia-driven flow network
in rodent brains strongly supports such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our results also revealed that solute concentrations are affected
by ventricular geometry changes.
In fact, a 9\% difference in volume
had a noticeable impact on solute transport,
suggesting that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF distribution and composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.

@@

The larval and adult zebrafish were raised under standard husbandry
conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 \textmu S with a
14/10 hr light/dark cycle.
From fertilization to 3 days post-fertilization (dpf), larvae were maintained in egg
water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW)
(1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were
\emph{ccdc103(dnaaf19)\textsuperscript{tn222a}}
(\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures},
\emph{Tg(5xUAS:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}
(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish larvae at the 2 dpf stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xUAS:Signal-Dendra2})
adult animals, as described by D'Gama~\emph{et al.}~\cite{DGama2025MotileBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

@@

\subsection{Wholemount zebrafish in vivo time-lapse imaging and photoconversion of Dendra2}
Two dpf larval zebrafish were anesthetized in 0.013\% MS222
(Ethyl 3-aminobenzoate methanesulfonate, Sigma) in AFW and
mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100),
to obtain a lateral view of the brain ventricles. After the
fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify.
After solidification, AFW containing 0.013\% MS222 was added,
and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss)
and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat)
at room temperature.

@@

\subsection{Intraventricular injection of microbeads and particle tracking}
We carried out particle tracking in a larval zebrafish injected with fluorescent beads,
as described in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

>>>>

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function. CSF
provides mechanical support for the brain and helps
distribute nutrients, neurotransmitters and metabolites throughout the
central nervous system. CSF flow is driven by several processes,
including the beating of motile cilia located on the
walls of the brain ventricles. Despite the physiological importance of
CSF, the underlying mechanisms of CSF flow and solute transport
in the brain ventricles remain to be comprehensively
resolved. This study analyzes and evaluates specifically
the role of motile cilia in CSF flow and transport. 
We developed finite element methods for modeling flow
and transport using the geometry of embryonic zebrafish
brain ventricles, for which we have detailed knowledge of
cilia properties and CSF motion.
The computational model is validated by in vivo experiments
that monitor transport of a photoconvertible protein secreted in the
brain ventricles. Our results show that while cilia contribute to
advection of large particles, diffusion plays a significant role in
the transport of small solutes. We also demonstrate how cilia location
and the geometry of the ventricular system impact solute
distribution. Altogether, this work presents a computational framework
that can be applied to other ventricular systems, together with new
concepts of how molecules are transported within the brain and its
ventricles.

@@

% I4: How do we address the problem and what do we find out?
In this study, we combined computational and experimental approaches to investigate
the role of motile cilia in brain ventricular CSF flow and transport.
We model the flow and transport by partial differential equations and solve
them numerically with finite element methods to compute flow patterns and
solute concentration dynamics. A geometry representing the embryonic zebrafish brain ventricles
was used, for which we have detailed knowledge of cilia properties and CSF motion velocities
and frequencies~\cite{Olstad2019CiliaryDevelopment}. 
Notably, this early developmental stage allows us to simplify our model
by omitting CSF secretion, which starts at later stages~\cite{Jeong2024TheZebrafish},
while still being well-conserved with mammals~\cite{DGama2025MotileBrain,
Olstad2019CiliaryDevelopment, Ringers2020Role, DGama2021Diversity, jurisch2020radial}. 
The computational model is validated against experimental data by monitoring in vivo
the dynamics of a photoconvertible fluorescent protein Dendra2
expressed within the brain ventricles.
Our results show that diffusion plays a major role in the transport of small solutes,
whereas cilia are paramount for the movement of larger particles.
We also found that the location of cilia, as well as changes to the ventricular geometry,
largely impact solute distribution.
In summary, our work presents a computational
framework that can be applied to other
ventricular systems in the future, together with new concepts
of how cilia distribute molecules and
particles within the brain ventricles.

@@

In vivo particle tracking reveals characteristic CSF flow patterns in
embryonic zebrafish brain ventricles
(\Crefpanels{fig:fig1}{a--c})~\cite{Olstad2019CiliaryDevelopment}.
We attempt to accurately represent these flow patterns with
computational fluid dynamics driven by cardiac pulsations and motile
cilia lining the surface of the brain ventricles.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish embryos expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein (\href{https://doi.org/10.5281/zenodo.15194757}{Supplementary Video S2}).
To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F(t)
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure3_compare_experiments_and_simulations.png}
    \caption{
    \textbf{a.} Relative change in fluorescence intensity $\Delta F(t)$ 
    in each region of interest (ROI)
    observed during experiments monitoring
    the movement of Dendra2 proteins photoconverted by laser targeting ROI 1. 
    Bold lines are mean values of the cohort ($n=8$) and shaded regions cover
    one standard deviation.
    \textbf{b.} Mean concentration $\cbar(t)$ in each ROI in a
    simulation of Dendra2 transport with
    the diffusion coefficient $D_3$.
    \textbf{c.} Time to $\Delta F(t)$ reaches a threshold value of 0.25 in
    one representative zebrafish embryo.
    \textbf{d.} Time to $\cbar(t)$ reaches a threshold value of 0.25 for
    a transport simulation with diffusion coefficient $D_3$.
    The posterior-most regions of the geometries are excluded,
    because $\Delta F$ and $\cbar$ never exceeded the threshold
    value of 0.25 in these regions. Scale bars 50 \textmu m.}
    \label{fig:fig3}
\end{figure}
% Final Delta F values
%0.6322650156762044 ROI 2
%0.7059449124175479 ROI 3
%0.6694500367160006 ROI 4
%0.28986700094045437 ROI 5
%0.18155055624229066 ROI 6
% Final mean c values ROIs 2-6: 0.80, 0.76, 0.86, 0.60 and 0.38

@@

We previously characterized three distinct ciliated cell lineages in
the embryonic zebrafish brain~\cite{Olstad2019CiliaryDevelopment,
DGama2025MotileBrain}.  To identify the effects of these lineages on
CSF flow and transport, we investigate region-specific cilia paralysis
with the computational model.  Three cilia-covered regions were
paralyzed separately: the anterior ventricle, and the dorsal and
ventral parts of the middle ventricle (\Crefpanels{fig:fig5}{a}).  To
simulate regional cilia paralysis, we remove the tangential traction exerted by
the cilia in the respective region.

@@

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow
patterns, and we previously identified large
interindividual variation in ventricular size and
CSF velocities between embryos at given developmental stages and throughout
development~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain, Jeong2024TheZebrafish}.
To assess how these
variations influence solute distribution, we consider four alternative
ventricular geometries: (i) constriction of the anterior-middle
interventricular duct (66 \% reduction in cross-sectional area);
(ii) constriction of the middle-posterior duct (33 \% reduction);
(iii) middle ventricle shrunk in the lateral direction (43 \% reduction);
and (iv) all three combined (\Crefpanels{fig:fig6}{a, b}).
In these four modified geometries, the total volume is reduced by
3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively, compared to the original
geometry. As a result of the geometry modifications,
the forces applied with the 
cilia tangential traction differ across the scenarios
(\Crefpanels{fig:fig6}{c}).

@@

\section{Discussion}
In this study, we have presented a computational model of CSF flow and
solute transport within brain ventricles.  We used the embryonic
zebrafish brain ventricles since their geometry and flow mechanisms are
well described, thereby allowing for validation of the simulation
results against experimental measurements. Our computational framework
can be adapted to other ventricular systems, once their flow parameters
are adequately resolved. Here, we assumed that
intraventricular CSF flow was governed by the Stokes equations. By
imposing tangential traction and normal pressure acting on the CSF, we
were able to reproduce flow features observed experimentally. In
particular, vortex-structured flow mediated by the cilia and pulsatile
flow mediated by the cardiac cycle emerged as key flow characteristics.

@@

Removing motile cilia delayed transport within
the middle ventricle and towards the anterior ventricle,
while the transport to the posterior ventricle was unaffected.
Photoconversion experiments with cilia mutant
zebrafish embryos (\emph{smh}) aligned with these observations.
These simulations and experiments further indicate that
transport from the middle to the posterior ventricle is
purely diffusive at this developmental stage,
while transport within the middle ventricle
depends on cilia motility.
Furthermore, our experimental results provide quantification
of the impact of cilia motility
on brain ventricular transport.
Interestingly, Yoshida~\emph{et al.} observed a drastic reduction
in CSF mixing in specific regions of the human brain
ventricles when simulating reduced cilia motility~\cite{Yoshida2022EffectVentricles},
an observation that compares well with our findings.

@@

We modeled motile cilia as a tangential
traction at the ventricular walls
while allowing the CSF to slip,
motivated by experimental observations of
flow fields in embryonic zebrafish brain ventricles in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The ciliated boundaries in our model are regionally homogeneous,
whereas in reality cilia populate the ventricle walls
heterogeneously~\cite{Olstad2019CiliaryDevelopment}. Heterogeneity
in cilia location, beating direction and beating frequency
enhances particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Beating directionality can affect transport,
and specific coordination of the ciliary beating is important for
flow and signaling processes in the brain ventricles
\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
for physiological mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet},
and in the nodal flow that establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry, Sawamoto2006NewBrain}.
Incorporating spatial heterogeneity in our model would
therefore presumably affect results, and could be
interesting to pursue in further work.

@@

\subsection{Outlook}
Altogether, we have shown that the flow
generated by ciliary motion in embryonic zebrafish brain ventricles
is of the scale of the geometry~\cite{Olstad2019CiliaryDevelopment}.
In contrast, one can expect that in large brains such as the
human brain, ciliary beating contributes only to flow proximal
to the ventricular walls\cite{Siyahhan2014FlowVentricles}.
Nevertheless, the cilia may still play an important role in regulating the
amount of substances, especially extracellular vesicles,
close to the ventricular walls,
similar to coral reefs where cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Notably, the discovery of a cilia-driven flow network
in rodent brains strongly supports such selective
brain ventricular transport~\cite{Faubel2016Cilia-basedVentricles}.
Our results also revealed that solute concentrations are affected
by ventricular geometry changes.
In fact, a 9\% difference in volume
had a noticeable impact on solute transport,
suggesting that pathological ventricular volume changes,
such as the enlargement observed in hydrocephalus,
could affect CSF distribution and composition.
Because CSF plays an important role in the regulation of solutes
within the ventricular system~\cite{yamada2021cerebrospinal},
understanding more of the interplay between
transport dynamics and ventricular morphology,
as well as the impact of
reduced mixing resulting from cilia-deficiency,
are important considerations for future work.

@@

The embryonic and adult zebrafish were raised under standard husbandry
conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 \textmu S with a
14/10 hr light/dark cycle.
From fertilization to 3 days post-fertilization (dpf), embryos were maintained in egg
water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW)
(1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were
\emph{ccdc103(dnaaf19)\textsuperscript{tn222a}}
(\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures},
\emph{Tg(5xUAS:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}
(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Wholemount in vivo live imaging and photoconversion
experiments were performed with zebrafish embryos at the 2 dpf stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xUAS:Signal-Dendra2})
adult animals, as described by D'Gama~\emph{et al.}~\cite{DGama2025MotileBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+})
or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

@@

\subsection{Wholemount zebrafish in vivo time-lapse imaging and photoconversion of Dendra2}
Two dpf embryonic zebrafish were anesthetized in 0.013\% MS222
(Ethyl 3-aminobenzoate methanesulfonate, Sigma) in AFW and
mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100),
to obtain a lateral view of the brain ventricles. After the
fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify.
After solidification, AFW containing 0.013\% MS222 was added,
and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss)
and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat)
at room temperature.

@@

\subsection{Intraventricular injection of microbeads and particle tracking}
We carried out particle tracking in an embryonic zebrafish injected with fluorescent beads,
as described in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.
